Future challenges for Research Software Engineering

### 2025-08-11 correspondence/rebuttal-1.md

> 2. The authors switch between the past and present tenses, which makes reading the article difficult. They sometimes address present tools, infrastructure and activities, and sometimes write as if future tools, infrastructure and activities have been made available or developed. For example, the "Achieved Goals" refer to goals achieved by 2035, but the tenses change throughout the boxes. "RSEng-relevant open educational resources..." Some resources are available and some are not; however, the style of writing makes it difficult to tell the difference. I would propose that the authors use tenses to clearly distinguish between past, present and future activities, tools and resources, and not write as if future activities have already been successful. Additionally, the authors write about their hopes, which seem off-topic for a report or opinion paper. I would suggest not writing about hopes at all. Future trend statements should be based on knowledge.

> 3. The authors claim to "present here the results of a workshop discussion at the deRSE25 conference in Karlsruhe". However, they also include their hopes, discussions and own expectations for the future RSE ecosystem (in Germany). This should be mentioned in the introduction and it should be made clear which statements have been put forward during the discussion and which have been added by the authors after the discussion. The authors mention this construct in section 3 ("In the previous sections, we have tried to list some of the trends that we foresee. The material presented here is based on the outputs of a workshop, but extended with the authors’ ideas."), which makes understanding the paper difficult; I would expect such a statement at the beginning. While I do not propose to re-structure the paper (e.g. report first, discussion second), I suggest to review the paper as to which statements have been reported in the discussion and which statement have been brought up by the authors.

We have put this sentence to the front, and kept it at the end as a reminder.

@@

> Another very minor point: the gender pronouns of the two personas ('Kim' and 'Kay') appear to swap at times; e.g. in Story 3 'Kay' is 'her' and in Story 4 'Kim' is 'him', while in Story 6 'Kay' is 'he' and 'Kim' is 'she')

### Suggestions

> If possible, the inclusion of a small table, figure, and/or a few lines of text summarising the distribution of key features of the workshop participants might be an interesting addition (e.g. level of academic/research education, an indication of their software engineering proficiency/professionalism, their age and/or career stage)... I would be curious as to the level of 'optimism' and expectations for change for the field that participants with different backgrounds might tend to have, although I feel that your paper has merit and interest as it stands, with no such changes needed.

Thank you for your suggestion. We made the workshop notes available on Zenodo and included the link into the paper. We definitely agree that more statisitics about the participants would be valuable, but we did not collect any such data. We also added a respective statement to the introduction.

### 2025-08-11 future_challenges_for_RSEng.tex
Update future_challenges_for_RSEng.tex
Co-authored-by: Gerasimos Chourdakis <gerasimos.chourdakis@ipvs.uni-stuttgart.de>

In this opinion paper, we try to imagine the state of our field
and community in the mid-term future, focusing on future challenges.
While future projections are common in industry~\cite{Gartner2025},
so far there is a projection lacking for the RSEng space.
We present here the results of a workshop discussion (notes content is archived here \cite{goth2025_WSPad}) at the
deRSE25 conference in Karlsruhe\footnote{deRSE25, session ``The End of {RSEng}?
Challenges and Risks for {RSEng}'': \url{https://events.hifis.net/event/1741/contributions/14026/}},
without claiming to represent the entire deRSE community.
The material presented here is based on the output of the workshop participants, but extended with the authors' ideas.
Since deRSE25 was co-located with SE25, the annual conference for the Software Engineering community in Germany, the discussion was also open to software engineers,
but a precise statistic of the workshop participants background was not taken.
The workshop hosts felt, that the majority of participants were from the RSE community.
Nevertheless, by disseminating the discussed ideas,
we aim to engage the community in creating a future worth living
and to advance the conversation about how Research Software Engineering should, or could, develop.

>>>>

In this opinion paper, we try to imagine the state of our field
and community in the mid-term future, focusing on future challenges.
While future projections are common in industry~\cite{Gartner2025},
so far there is a projection lacking for the RSEng space.
We present here the results of a workshop discussion (notes content is archived here \cite{goth2025_WSPad}) at the
deRSE25 conference in Karlsruhe\footnote{deRSE25, session ``The End of {RSEng}?
Challenges and Risks for {RSEng}'': \url{https://events.hifis.net/event/1741/contributions/14026/}},
without claiming to represent the entire deRSE community.
The material presented here is based on the output of the workshop participants, but extended with the authors' ideas.
Since deRSE25 was co-located with SE25, the annual conference for the Software Engineering community in Germany, the discussion was also open to software engineers,
but no list or profile of the participants is available.
The workshop hosts felt, that the majority of participants were from the RSE community.
Nevertheless, by disseminating the discussed ideas,
we aim to engage the community in creating a future worth living
and to advance the conversation about how Research Software Engineering should, or could, develop.
